---
title: "STS_simple_data_workflow"
author: "Holly Doerr"
date: "2023-07-19"
output: html_document
---

This is our workflow for manipulating the simple sand tiger shark tag and reciever data. 

First, we will load our reciever data files and assign individual dataframes. 

```{r read files, eval=FALSE, message = FALSE, warning = FALSE}
#You can set your working directory to make sure your code is pulling the files you want.
setwd("C:/Users/hdoer/Documents/STS_data/6224302/6224302")

#install.packages("dplyr")
#install.packages("tidyverse")
library(tidyverse)

#load the data file and assign as a dataframe: 
data_2020 <- read.csv("uncst_matched_detections_2020.csv") #read in data

data_2021 <- read.csv("uncst_matched_detections_2021.csv") #read in data

data_2022 <- read.csv("uncst_matched_detections_2022.csv") #read in data

data_2023 <- read.csv("uncst_matched_detections_2023.csv") #read in data


head(data_2020)
head(data_2021)
head(data_2022)
head(data_2023)
```
Now that we have loaded our individual reciever data files, let's start by combining the files to make one big master dataframe using our bind_rows() function - this makes it easier to index the data as a whole downstream. 

```{r}
master_data <- bind_rows(data_2020, data_2021, data_2022, data_2023, .id=NULL)

head(master_data)

write.csv(master_data, 'master_UNCST_matched_detections_data_2020_2023.csv')
```

Now we have our master file! Before we jump into master file queries, let's look at some data per year: 

```{r}
#number of sharks:
sharks_2020 <- data_2020 %>% group_by(tagname) %>% tally()#this tallies how many times each tagname appears, aka how many times each individual shark pinged, and outputs to a new table
sharks_2020 <- sharks_2020 %>% arrange(desc(n)) #this rearranges the tallied data to be in descending order to easily determine min and max pings

sharks_2021 <- data_2021 %>% group_by(tagname) %>% tally()
sharks_2021 <- sharks_2021 %>% arrange(desc(n))

sharks_2022 <- data_2022 %>% group_by(tagname) %>% tally()
sharks_2022 <- sharks_2022 %>% arrange(desc(n))

sharks_2023 <- data_2023 %>% group_by(tagname) %>% tally()
sharks_2023 <- sharks_2023 %>% arrange(desc(n))


#number of pinged locations: 

locations_2020 <- data_2020 %>% group_by(station) %>% tally() #this tallies how many times each location is associated with an individual shark's ping and outputs to a new table. 
locations_2020 <- locations_2020 %>% arrange(desc(n)) #this rearranges the tallies to be in descending order 

locations_2021 <- data_2021 %>% group_by(station) %>% tally()
locations_2021 <- locations_2021 %>% arrange(desc(n))

locations_2022 <- data_2022 %>% group_by(station) %>% tally()
locations_2022 <- locations_2022 %>% arrange(desc(n))

locations_2023 <- data_2023 %>% group_by(station) %>% tally()
locations_2023 <- locations_2023 %>% arrange(desc(n))
```

Now let's look at things for the total set: 
```{r}
sharks_total <- master_data %>% group_by(tagname) %>% tally()
sharks_total <- sharks_total %>% arrange(desc(n))

locations_total <- master_data %>% group_by(station) %>% tally()
locations_total <- locations_total %>% arrange(desc(n))
```

So we've crunched some simple numbers, now let's see if we can visualize these data in stacked bar graphs!

First, we can simplify our data sets to create new data frames that contain only the data we need for the graphs.
```{r}
#we select just the tagname and station columns for each of our master, 2020, 2021, 2022 and 2023 files.

master_data_specific <- select(master_data, tagname, station)
head(master_data_specific)

data_2020_specific <- select(data_2020, tagname, station)
head(data_2020_specific)

data_2021_specific <- select(data_2021, tagname, station)
head(data_2021_specific)

data_2022_specific <- select(data_2022, tagname, station)
head(data_2022_specific)

data_2023_specific <- select(data_2023, tagname, station)
head(data_2023_specific)
```


```{r}
# library
install.packages("viridis")
library(ggplot2)
library(viridis)
#install.packages("hrbrthemes")
#library(hrbrthemes)
```
Next we will plot our data in stacked bar graphs to visualize in a different way!

```{r}
 
# Master plot stacked


Total_data <- ggplot(master_data_specific, aes(fill=station, x= tagname)) + 
    geom_bar() +
    #scale_y_continuous(labels = scales::percent) +
    #scale_fill_viridis(discrete = T) +
    ggtitle("Sand Tiger shark tracker pings along the East Coast from 2020 - 2023") +
    theme(legend.key.size = unit(0.5, units = "mm"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 3),
        plot.title = element_text(size=10),
        axis.title = element_text(size=14,face="bold"),
        #change x text size
        axis.text.x = element_text(angle = 90, size = 10))+
    xlab("")

Total_data

ggsave(filename = "master_graph.png", plot = Total_data)
```

```{r}
# 2020 plot stacked

data_2020_graph <- ggplot(data_2020_specific, aes(fill=station, x= tagname)) + 
    geom_bar() +
    #scale_y_continuous(labels = scales::percent) +
    #scale_fill_viridis(discrete = T) +
    ggtitle("Sand Tiger shark tracker pings along the East Coast from 2020 - 2023") +
    theme(legend.position = "none", 
        #legend.key.size = unit(0.5, units = "mm"),
        #legend.title = element_text(size = 10),
        #legend.text = element_text(size = 3),
        plot.title = element_text(size=12, face="bold"),
        axis.title = element_text(size=10,face="bold"),
        #change x text size
        axis.text.x = element_text(angle = 90, size = 10))+
    xlab("")

data_2020_graph

ggsave(filename = "2020_data_graph.png", plot = data_2020_graph)

library(cowplot)
library(grid)
library(gridExtra)

#Do this to save the legend separately. Must graph plot with legend first, then go back and resave image without legend.
legend <- cowplot::get_legend(data_2020_graph)
grid.newpage()
grid.draw(legend)

ggsave(filename = "2020_data_legend.png", plot = legend, width = 10, height= 7)

```

```{r}

#2021 plot stacked

data_2021_graph <- ggplot(data_2021_specific, aes(fill=station, x= tagname)) + 
    geom_bar() +
    #scale_y_continuous(labels = scales::percent) +
    #scale_fill_viridis(discrete = T) +
    ggtitle("Sand Tiger shark tracker pings along the East Coast during 2021") +
    theme(legend.position = "none", 
        #legend.key.size = unit(0.5, units = "mm"),
        #legend.title = element_text(size = 10),
        #legend.text = element_text(size = 3),
        plot.title = element_text(size=12, face="bold"),
        axis.title = element_text(size=10,face="bold"),
        #change x text size
        axis.text.x = element_text(angle = 90, size = 10))+
    xlab("")

data_2021_graph

ggsave(filename = "2021_data_graph.png", plot = data_2021_graph)

library(cowplot)
library(grid)
library(gridExtra)

#Do this to save the legend separately. Must graph plot with legend first, then go back and resave image without legend.
legend <- cowplot::get_legend(data_2021_graph)
grid.newpage()
grid.draw(legend)

ggsave(filename = "2021_data_legend.png", plot = legend, width = 14, height= 7)
```

```{r}
#2022 plot stacked

data_2022_graph <- ggplot(data_2022_specific, aes(fill=station, x= tagname)) + 
    geom_bar() +
    #scale_y_continuous(breaks = seq(0, 1700, by=50), limits=c(0,1700)) +
    #scale_fill_viridis(discrete = T) +
    ggtitle("Sand Tiger shark tracker pings along the East Coast during 2021") +
    theme(legend.position = "none", 
        #legend.key.size = unit(0.5, units = "mm"),
        #legend.title = element_text(size = 10),
        #legend.text = element_text(size = 3),
        plot.title = element_text(size=12, face="bold"),
        axis.title = element_text(size=10,face="bold"),
        #change x text size
        axis.text.x = element_text(angle = 90, size = 10))+
    xlab("")

data_2022_graph

ggsave(filename = "2022_data_graph.png", plot = data_2022_graph)

library(cowplot)
library(grid)
library(gridExtra)

#Do this to save the legend separately. Must graph plot with legend first, then go back and resave image without legend.
legend <- cowplot::get_legend(data_2022_graph)
grid.newpage()
grid.draw(legend)

ggsave(filename = "2022_data_legend.png", plot = legend, width = 10, height= 7)
```